---
title: "SoilGrids250 ground overlays"
author: "D G Rossiter"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_align: center
    fig_height: 6
    fig_width: 6
    number_section: yes
    theme: spacelab
    df_print: paged
    toc: yes
    toc_float: yes
---


# Objective

The aim of this script is to facilitate evaluation of [ISRIC-World Soil Information](https://www.isric.org)'s SoilGrids250 gridded maps of soil properties. Here we create ground overlay KML files for display in Google Earth. 
This should facilitate a qualitative evaluation of the pattern of SoilGrids250 predictions, because they can be visualized on the land cover and terrain.

If the overlay is specified for the continental USA (48 States), the overlay can be displayed along with [SoilWeb Apps](https://casoilresource.lawr.ucdavis.edu/soilweb-apps/), including SoilWeb Earth which displays SSURGO (detailed) or NATSGO (generalized) soil map units in Google Earth. The USA has a long-established soil survey, and most of the lower 48 States have been mapped in detail with extensive field investigation and compiled into polygon maps.

# Usage

This documents has two parts. The first is a step-by-step explanation with a single Area of Interest (AOI) which can be adjusted manually [in this section](#aoi), property, depth and quantile, which can be adjusted manually [in this section](#voi). The second is a self-contained [function](#function) to carry out the analysis for a user-selected set of properties, depths and quantiles for a single AOI and save all the results in the file system; this is [here](#compute).  When compiling this document with `knitr`, sample calls to the function are suppressed with the `eval=FALSE` markdown tag.

_Usage (1)_: To create one overlay step-by-step, specify an area of interest (AOI) [in this section](#aoi), and a property, depth and quantile  [in this section](#voi). Then either run directly or knit to HTML. The ground overlay (as PNG graphics file) and KML specification will be placed in the directory structure based on the AOI, property, depth and quantile. Open the KML in Google Earth and adjust transparency as desired.

_Usage (2)_: To create one more more overlays with the function, first run [this code](#function) to define the function `make.ground.overlay` in your workspace, and then call it, as shown with examples in [this section](#compute).

# Setup

Here is the base URL for local storage of the downloaded SoilGrids250 coverage and the PNG/KML built from this. Adjust to your system.


## Storage

Base directories for import from the SoilGrids server, and for writing the ground overlays. Note that the import must have been already done with script `SoilGrids250_import.Rmd`, which uses the same directory structure.

```{r}
base.dir <- "/Volumes/Pythagoras/ds"
base.dir.sg.kml <- paste0(base.dir, "/SoilGrids250_kml") 
base.dir.sg.import <- paste0(base.dir, "/DSM_export/SoilGrids250")
```

The produced KML will be stored under `r  print(base.dir.sg.kml)`.

## Packages

```{r message=FALSE}
options("rgdal_show_exportToProj4_warnings"="none") 
library(XML)            # various XML utilities
library(terra)          # raster data, replaces `raster`
library(raster)         # some functions still not in `terra`
library(rgdal)          # GDAL access from R
library(gdalUtils)      # wrappers for GDAL utility programs that could be
                        #  called from the command line
library(sf)             # spatial data types 
library(sp)             # older representation of spatial data types
library(maptools)       # create KML overlays, among others
```

## Set variable of interest {#voi}

Define the variables for the SoiLGrids250 soil property and layer of interest. See [here](https://www.isric.org/explore/soilgrids/faq-soilgrids#What_do_the_filename_codes_mean) for the naming conventions and units of measure.

`Q0.05` - 5% quantile from the Quantile Random Forest (QRF);
`Q0.5` - median of the distribution from the QRF -- note *not* `Q0.50`;
`mean` - mean of the distribution;
`Q0.95` - 95% quantile from the QRF.

Here are lists of the legal properties, quantiles, and depth slices:

```{r}
voi.list <- c("bdod", "cec", "cfvo", "clay", 
              "nitrogen", "phh2o", "sand", "silt",
              "soc", "ocd", "ocs")
quantile.list <- c("Q0.05", "Q0.5", "Q0.95", "mean")
depth.list <- c("0-5", "5-15", "15-30", "30-60", "60-100", "100-200")
```

Set the property, depth and quantile:

```{r}
voi <- "phh2o"   # variable of interest
depth <- "0-5cm" # depth slice
quantile <- "mean"  # quantile or mean
```

## Define Area of Interest (AOI) {#aoi}

Specify the lower-right corner using WGS84 geographic coordinates and the longitude/latitude extent, in degrees.

```{r}
tile.lrc <- c(-76, 42) # lower-right corner: central NY
d.long <- 1; d.lat <- 1
```

Compute the bounding box:

```{r}
.tile.ulc <- c(tile.lrc[1]-d.long, tile.lrc[2]+d.lat) # upper-left corner
.m <- matrix(c(.tile.ulc[1],tile.lrc[1],  #ulc
              .tile.ulc[2], tile.lrc[2]  #lrc
              ), nrow=2)
bb.ll <- st_sfc(st_multipoint(.m))
st_crs(bb.ll) <- 4326
```

A prefix for directories, to keep AOI results separate.

```{r}
AOI.dir.prefix <- paste0("lat", tile.lrc[2], .tile.ulc[2],
                         "_lon", .tile.ulc[1], tile.lrc[1])
```


## Read into R

We use the `terra` package to read the GeoTIFF. This must have been previously imported with script `SoilGrids250_import.Rmd`, into this directory structure.

```{r get-sg}
voi_layer <- paste(voi, depth, quantile, sep="_")
file.in <- paste0(base.dir.sg.import, "/", 
                  AOI.dir.prefix, "/", 
                  voi, "/",
                  quantile, "/",
                  depth, "/",
                  voi_layer, '.tif')
r.sg <- terra::rast(file.in)
```

The 0 values are masked urban and water. Convert these to `NA`.

```{r zero-to-na}
rcl <- matrix(c(0, NA), nrow=1)
r.sg <- terra::classify(r.sg, rcl, include.lowest=TRUE)
print(r.sg)
summary(r.sg)
plot(r.sg, col=heat.colors(256),
     main=voi_layer)
```

# Creating a ground overlay and KML specification

Google Earth uses [KML](https://developers.google.com/kml/documentation/kml_tut) (``Keyhole Markup Language'') to specify how to display geographic data. This is a tag-based structure with nested elements and attributes and is based on the XML standard, but with tags specific to geographic data.

So to display coverages created in R, there are five steps:

1. Create a map to export;
2. Assign projection information to the map;
3. Transform to geographic co\"{o}rdinates (Long/Lat) on the WGS84 datum, as required by Google Earth;
4. Export as a KML file;
5. Open the KML file in Google Earth.

Here we already have the map to export, with known projection, see previous section.

KML files can be point-based (placemarks), but here we need to produce so-called *ground overlays* in the Portable Network Graphics (PNG) graphics format, with accompanying information to tell Google Earth where in the world to place this.

This task is made possible by methods from the ``maptools` package, in particular the `GESpatialGrid` method to define the size and position of a PNG image overlay in Google Earth, and the `kmlOverlay` method write the KML file including this PNG image overlay.


## Convert to WGS84 raster

Google Earth uses exclusively the WGS84 CRS.

```{r}
bbox(r.sg)
res(r.sg)
r.sg.84 <- terra::project(r.sg, "+proj=longlat +datum=WGS84",
                          method="bilinear")
bbox(r.sg.84)
res(r.sg.84)
plot(r.sg.84, col=heat.colors(256),
     main=voi_layer)
```

The bounding box is somewhat larger than the requested tile, because converting to Homolosine gave more extreme values at opposite corners, and this was the portion of SoilGrids250 that was downloaded.

## Create a regular grid

The `maptools` package requires an object to be of type `GE_SpatialGrid`.

A `GE_SpatialGrid` has a triple purpose:

1. to contain the graphics file (PNG) to be displayed in Google Earth;
2. to give information for setting up the PNG graphics device in R, i.e. the number of pixels;
3. to provide display information for Google Earth, i.e. the geographic location (bounding box) and the aspect ratio.

The function requires a `Spatial*` object from package `sp`. We must first convert the `terra::SpatRaster` to a `raster::RasterLayer` from the older `raster` package (which `terra` is replacing), and then cast to an `sp` object.

```{r create.spdf}
class(r.sg.84)
tmp <- raster::raster(r.sg.84)
class(tmp)
r.sp.84 <- as(tmp, "SpatialPixelsDataFrame")
class(r.sp.84)
# spplot(r.sp.84, col.regions=heat.colors(256))
```

Now make the `GE_SpatialGrid`, changing the default pixel size to show each prediction pixel.

```{r warning=FALSE}
# default maxPixels=600, we want more detail
GRD <- maptools::GE_SpatialGrid(r.sp.84, maxPixels=4800)
str(GRD)
```

## Export to PNG

This is the ground overlay.

```{r dest.dir.local}
# local storage
dest.dir.sg <-  paste0(base.dir.sg.kml, "/",
                       AOI.dir.prefix, "/",
                       voi, "/", 
                       quantile, "/", 
                       depth)
if (!dir.exists(dest.dir.sg)) {
  dir.create(dest.dir.sg, recursive = TRUE)
}
```

```{r}
filename.png <- paste0(dest.dir.sg,"/", voi_layer, '.png')
png(file=filename.png, width=GRD$width, height=GRD$height,
    bg="transparent")
par(mar=c(0,0,0,0), xaxs="i", yaxs="i")
image(r.sp.84, xlim=GRD$xlim, ylim=GRD$ylim, 
     col=heat.colors(256, alpha=1),  # maximum colour resolution for PNG, opaque
     setParUsrBB=TRUE)  # use our bounding box
dev.off()
```

# create a colour scale to add to the KML display

```{r color.bar}
.min <- floor(min(r.sp.84@data))
.max <- ceiling(max(r.sp.84@data))
.ramp <- heat.colors(256, alpha=1)
.scale <-  (length(.ramp)-1)/(.max-.min)
.ticks <- seq(.min, .max, len = 11)
.filename <- paste0(dest.dir.sg, "/", 
                    voi_layer, '_colorbar.png')
png(.filename, width=GRD$width/16, 
    height=GRD$height/6,
    pointsize = 14,
    type = "quartz",
    units = "px")
plot(c(0,30), c(.min,.max), type='n', bty='n', 
     xaxt='n', xlab='', yaxt='n', ylab='', main="")
axis(2, .ticks, las=1)
for (.i in 1:(length(.ramp)-1)) {
  .y = (.i-1)/.scale + .min
  rect(0,.y,30,.y+1/.scale, col = .ramp[.i], border = NA)
}
dev.off()
```


## Create KML

This specifies the ground overlay to display (a separate file from the KML) and its position. The filename is identical to that for the PNG, but with a different extension.

```{r}
filename.kml <- paste0(dest.dir.sg,"/", voi_layer, '.kml')
maptools::kmlOverlay(GRD, 
                     kmlfile = filename.kml,
                     imagefile = filename.png,
                     name=paste0("SoilGrids250", voi_layer)
                     )
```


Now this file can be opened in Google Earth.

# A function to automate this process {#function}

Here we take the essentials of the step-by-step approach above, and make a function which will carry out the analysis with four argments:

* Property of interest
* Depth interval
* Quantile or mean
* lower-right corner (default: central NL) and size (default $1 \times 1^\circ$) of the tile , in degrees

This function has the side effect of saving the ground overlay (PNG and KML specification) under the appropriate subdirectories.

```{r define.function, warning=FALSE}
make.ground.overlay <- function(voi, depth, quantile, tile.lrc=c(52, 6), d.long=1, d.lat=1) {
  
  # local storage
  base.dir <- "/Volumes/Pythagoras/ds"
  base.dir.sg.kml <- paste0(base.dir, "/SoilGrids250_kml") 
  base.dir.sg.import <- paste0(base.dir, "/DSM_export/SoilGrids250")

  # packages   
  library(XML)            # various XML utilities
  library(terra)          # raster data, replaces `raster`
  library(raster)         # some functions still not in `terra`
  library(rgdal)          # GDAL access from R
  library(gdalUtils)      # wrappers for GDAL utility programs that could be
  #  called from the command line
  library(sf)             # spatial data types 
  library(sp)             # older representation of spatial data types
  library(maptools)       # create KML overlays, among others
  
  # check arguments
  
  # setup
  voi_layer <- paste(voi, depth, quantile, sep="_")
  # AOI
  .tile.ulc <- c(tile.lrc[1]-d.long, tile.lrc[2]+d.lat) # upper-left corner
  .m <- matrix(c(.tile.ulc[1],tile.lrc[1],  #ulc
                 .tile.ulc[2], tile.lrc[2]  #lrc
  ), nrow=2)
  bb.ll <- st_sfc(st_multipoint(.m))
  st_crs(bb.ll) <- 4326
  .crs.igh <- '+proj=igh +lat_0=0 +lon_0=0 +datum=WGS84 +units=m +no_defs'
  .bb.igh <- st_transform(bb.ll, .crs.igh)
  .bb.igh <- st_coordinates(.bb.igh)[,1:2]
  AOI.dir.prefix <- paste0("lat", tile.lrc[2], .tile.ulc[2],
                           "_lon", .tile.ulc[1], tile.lrc[1])
  
  # local storage for imported tile
  voi_layer <- paste(voi, depth, quantile, sep="_")
  file.in <- paste0(base.dir.sg.import, "/", 
                    AOI.dir.prefix, "/", 
                    voi, "/",
                    quantile, "/",
                    depth, "/",
                    voi_layer, '.tif')
  
  ## check that this SG tile has been imported
  if (!file.exists(file.in)) {
    print("SoilGrids250 tile has not been imported")
    return(NULL)
  }
  
  ## read SoilGrids tile into R, replace 0 with NA
  r.sg <- terra::rast(file.in)
  r.sg <- classify(r.sg, matrix(c(0, NA), nrow=1), include.lowest=TRUE)
  
  # resameple to WGS84
  r.sg.84 <- terra::project(r.sg, "+proj=longlat +datum=WGS84",
                            method="bilinear")
  tmp <- raster::raster(r.sg.84)
  r.sp.84 <- as(tmp, "SpatialPixelsDataFrame")
  # grid on which to make the PNG
  GRD <- maptools::GE_SpatialGrid(r.sp.84, maxPixels=4800) # high resolution
  
  # local storage for ground overlay
  dest.dir.sg <-  paste0(base.dir.sg.kml, "/",
                         AOI.dir.prefix, "/",
                         voi, "/", 
                         quantile, "/", 
                         depth)
  if (!dir.exists(dest.dir.sg)) {
    dir.create(dest.dir.sg, recursive = TRUE)
  }
  
  # write the PNG
  filename.png <- paste0(dest.dir.sg,"/", voi_layer, '.png')
  png(file=filename.png, width=GRD$width, height=GRD$height,
      bg="transparent")
  par(mar=c(0,0,0,0), xaxs="i", yaxs="i")
  image(r.sp.84, xlim=GRD$xlim, ylim=GRD$ylim, 
        col=heat.colors(256, alpha=1),
        setParUsrBB=TRUE)
  dev.off()

  # write the KML, referring to the PNG
  filename.kml <- paste0(dest.dir.sg,"/", voi_layer, '.kml')
  maptools::kmlOverlay(GRD, 
                       kmlfile = filename.kml,
                       imagefile = filename.png,
                       name=paste0("SoilGrids250", voi_layer)
  )
}
```

# Calling the function {#compute}

Here are some examples. 

A single property, depth, quantile, $1 \times 1^\circ$ AOI:

```{r eval=FALSE}
make.ground.overlay("clay", "30-60cm", "mean", c(-76, 42), 1, 1)
```

The 5% and 95% quantiles of two properties at two depth slices, $1 \times 1^\circ$ AOI:

```{r eval=FALSE}
for (voi in c("clay", "phh2o")) {
   for (depth in paste0(c("0-5", "30-60"),"cm")) {
      for (quantile in c("Q0.05", "Q0.95")) {
         print(paste("Making overlay for property", voi, "; depth", depth, "; quantile", quantile))
         make.ground.overlay(voi, depth, quantile, c(-76, 42), 1, 1)
      }
   }
}
```
