---
title: "GlobalSoilMap USA v0.5 -- Import for PSM comparisons"
author:
  - "D G Rossiter"
  - "d.g.rossiter@cornell.edu"
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
params:
   lrc_long: -76
   lrc_lat: 42 
   size: 1
   quantile.n: NA
   voi.n: 4
   depth.n: 4
output:
  html_document:
    fig_align: center
    fig_height: 6
    fig_width: 6
    fig_caption: false
    number_section: yes
    theme: spacelab
    df_print: paged
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      purl=FALSE, 
                      fig.align = 'center')
knitr::opts_chunk$set(cache.extra = R.version.string, comment="")
```

# Introduction

[Global Soil Map v0.5](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/research/?cid=nrcseprd1321715), further abbreviated as `GSMv0.5` is the USA's contribution to the `GlobalSoilMap.net` project. It is based on the Soil Survey Geographic Database (SSURGO), at nominal 100 m grid resolution.

GSMv0.5 is partially explained in Hempel, J., Libohova, Z., Odgers, N., Thompson, J., Smith, C. A., & Lelyk, G. (2012). Versioning of GlobalSoilMap.net raster property maps for the North American Node. In Digital Soil Assessments and Beyond: Proceedings of the Fifth Global Workshop on Digital Soil Mapping 2012, Sydney, Australia (pp. 429–433). CRC Press. https://doi.org/10.13140/2.1.4670.2404.  A promised 1.0+ is supposed to use a DSM approach from NCSS-KSSL point observations, i.e., a conventional DSM approach.

This script creates a tile for a property and depth slice, over a Area of Interest delimited by geographic coordinates, that can then be compared with other PSM products. The original Coordinate Reference System (EPSG 4326, WGS84 long/lat) and grid resolution (3") are not changed.

To use this script:

1. Ajust the [directory structure](#dirs) to your system

2. [Select a property](#prop) and [select a depth slice](#depth), using the YAML header or by knitting with parameters.

3. [Select an Area of Interest](#aoi), using the YAML header or by knitting with parameters.

4. Either compile to HTML or PDF ("knit"), or "Run All" within R Markdown.

5. The processed tile will be in the directory structure, in a [subdirectory named for the AOI](#save).

# Packages and Drivers

```{r pack}
library(sf)
library(rgdal)
library(terra) 
library(tidyverse)
``` 

Check that the system has a driver for OpenFile Geodatabases:

```{r drive}
ogr.d <- ogrDrivers()
ogr.d[ix <- which(ogr.d$name=="OpenFileGDB"), ]
```


# Directories {#dirs}

One directory is used for the large original files (all-CONUS grid and all-map unit database), and another for the extracted tiles. Set these to areas on your own system.

```{r}
base.dir.gsm05 <- "/Volumes/Pythagoras/ds/DSM_export/GSM_USA/"
base.dir.gsm05.import <- "/Volumes/Pythagoras/ds/DSM_import/GSM_USA/"
```

# Files

These must be downloaded prior to running this script:

* [URL for all files](https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/):  https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/

* [Database](https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/GSM_34LRH_20180606094754.gdb.7z):  https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/GSM_34LRH_20180606094754.gdb.7z

Download and then uncompress into a directory named `GSM_34LRH_20180606094754.gdb`, under `base.dir.gsm05.import` (see above). This is an ESRI Open File Geodatabase with 43 files, total size 1.8 GB.

There are 1758 attribute fields. These are grouped into 26 'entities'. The present 'entity' describes the whole table. The next entity is labelled 'basic attributes' to collect 12 attributes that describe the map unit but do not represent soil properties. The remaining entities each describe a soil property and its method of analysis. The definitions related to the soil property and the links to NRCS metadata are given at the entity level, and the specific distinctions among data fields are given at the attribute level.

Each attribute that varies by depth was coded for 8 depth zones given here with their labels: 0-5 cm (000_005), 5-15 cm (005_015), 15-30 cm (015_030), 30-60 cm (030_060), 60-100 cm (060_100), 100-200 cm (100_200), the total profile (000_999), and for the plant exploitable (effective) depth of the soil (rootzone),

This version of the table also contains intermediate values used in the calculation of the primary GlobalSoilMap variables, which are not required by the specifications, such as the thickness, volume, and mass of soil which was used for weighting when computing weighted averages across horizons. It also includes metadata variables, such as the sum of the component percentages (`mu_sum_comppct_r_fix_`) which were used in the calculation, where a high value indicates that a large proportion of the land area is included in the calculated result, and a low value indicates that only a small proportion of the land area contributed to the result.

* [Download the grid](https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/mukeyint_SaS_3as.7z): 
 https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/mukeyint_SaS_3as.7z

* Uncompress into a directory named `mukeyint_SaS_3as`, under `base.dir.gsm05.import` (see above). This has a GeoTIFF and related metadata files; total size 9.8 GB.
<br>
The NoData values are 255 for 8-bit unsigned integers, 65535 for 16-bit unsigned integers, 4294967295 for 32-bit unsigned integers, and -3.39999995214436E+38 for 32-bit floating point numbers.

* [Download the metadata](https://edcftp.cr.usgs.gov/project/soil/GlobalSoilMap/T_GSM_20180430162220_metadata.xml) written by Norman Bliss. 
<br>
Download and open in a web browser that can interpret XML.


# Attribute database

Information:

```{r find.db}
gdb.name <- paste0(base.dir.gsm05.import, "GSM_34LRH_20180606094754.gdb")
file.info(gdb.name)[c("size", "mtime")]
(db.layers <- ogrListLayers(gdb.name))
```

Read and examine. 

This takes several minutes, so if it is already in memory (e.g., if you run interactively) do not re-read.

```{r read.db}
if (!exists("db")) {
  cat("Loading attribute database...")
  system.time(db <- sf::st_read(gdb.name))
} else {
  cat("Using previously-loaded attribute database.")
}
```

```{r info.db}
class(db)

table(db$mukind)  # some have no map unit kind!
head(db$muname)
head(db$mukey)
```

All features are NULL, because this is a large data table, not a map.

Example of property names, for clay. See Metadata (XML) for others.

```{r prop.names}
ix <- grep("clay", names(db))
sort(names(db)[ix])
```

Explanation from the metadata:

"In this group of attributes, there are 3 range_point cases (low: `_l`, representative value: `_r`, and high: `_h`). There are four principal variables: three intermediate variables for the calculation of the thickness (`mu_thick_cm3_cm2_claytotal_`), the volume of soil fines (`mu_vol_lt2mm_cm3_cm2_claytotal_`), and the mass of soil fines (`mu_mass_lt2mm_g_cm2_claytotal_`), and the *primary clay fraction variable* (`mu_claytotal_`) for each of the 8 depth_zone cases, as well as a `sum_comppct_` variable, for a subtotal of 4 * 8 + 1 = 33 attributes for each range_point case. There are two additional range point cases (low of the low: `_ll` and high of the high: `_hh`) with just the principal variables, and no additional thickness, volume, mass, or sum_comppct variables, for a subtotal of 8 attributes (by depth zone) for each additional range point case. Thus, there are 3 * 33 + 2 * 8 = 115 attributes in this group."

These are of the form:

* `mu_` : represent the entire map unit;
* `vol_lt2mm_`, `mass_lt2mm_`, `thick_`, `*total_` where `*` is the property; note `lt2mm` means $<2$mm particle diameter;
* `r_`, `h_`, `l_`, `hh_`, `ll_` what the value represents, as explained above;
* units of measure, e.g. `g_kg`, `g_gF`, `g_cm2`, `l_g`, `cm3_cm2` where the second is the denominator; note `_gF` is grams of fines;   
* depth slice in cm, e.g, `000_005`, `rootzone`; note `000_999` is over the whole profile.

`thick` is the thickness of the soil in the depth zone used in the calculation, computed as a weighted average of the components.

A special case are the component percentages, of the form:

```{r prop.props}
ix <- grep("mu_sum_comppct_r_fix_claytotal_[r,l,h]", names(db))
names(db)[ix]
```

* `mu_sum_comppct_r_fix_*total_{r, l, h}` where `*` is the property; this gives the proportion of the map unit from which this value is taken.

```{r comppct}
db$mu_sum_comppct_r_fix_claytotal_r[16:20]
db$mu_sum_comppct_r_fix_claytotal_l[16:20]
db$mu_sum_comppct_r_fix_claytotal_h[16:20]
```

These should be (?) the same, since the three numbers are taken from the dominant component (?).

For comparison with PSM predictions we want the `total` representative values, for clay by weight, i.e., in `_g_kgF`, i.e., %%.

## Uncertainty

SSURGO and STATSGO2 have a "low", "representative", and "high" value for most quantitative attributes in the component horizon (`chorizon`) attribute table. The "representative" values (with `_r` in the attribute name) are the primary attributes intended to characterize soils for most users. Where the "low" (`_l`) and "high" (`_h`) values are given, they are intended to help characterize the uncertainty in the representative values. An additional method of characterizing uncertainty is given by the "low of the low" (`_ll`) and "high of the high" values (`_hh`) where "`_ll`" represents the minimum value of the "low" across all components, and "`_hh`" represents the maximum value of the "high" across all components. 

# Parameters

Parameters for this run:

```{r}
print(paste("lrc_long:", params$lrc_long, "; lrc_lat:", params$lrc_lat, "; size:", params$size))
print(paste("voi.n:", params$voi.n, "; depth.n:", params$depth.n))
```

## Property of interest {#prop}

The following properties can be compared to SoilGrids250 and other PSM products. We need to specify the long GSM names because of the units of measure.

```{r voi.list}
voi.list.sg <- c("clay", "silt", "sand", "phh2o", "cec", "soc", "bdod", "cfvo")
voi.list.gsm <- c("claytotal_r_g_kg", "silttotal_r_g_kg", "sandtotal_r_g_kg",
                  "ph1to1h2o_r_ions_pHx10", "ecec_r_cmolc_kg", "soc_r_mr_g_gF", 
                  "dbthirdbar_lt2mm_r_g_cm3", "gravel_r_vol_ratio_m3_m3")
```

*Select a property* by its position in the list, and make a full name from it:

```{r voi}
(voi.name <- paste0("mu_", voi.list.gsm[params$voi.n], "_"))
```

### Representative values

Get the representative values for the selected property and _replace NoData values_ used in the database with R `NA`. The NoData values vary by data type. Also include the map unit key, for joining with the grid.

"Data are within the stated data ranges, except for the NoData values. The NoData values are 255 for 8-bit unsigned integers, 65535 for 16-bit unsigned integers, 4294967295 for 32-bit unsigned integers, and -3.4e38 for 32-bit floating point numbers." Actually the 32-bit floating point value is -3.39999995214436E+38, it's most reliable to get the exact value as the minimum of the corresponding value.

```{r}
dim(db)
head(names(db))
db.attr <- dplyr::select(db, matches(c("mukeyint", voi.name)))
names(db.attr)
dim(db.attr)
# the 2nd attribute is typical of all depths
# set the NoData values
na.val <- ifelse(is.integer(db.attr[,2]), 2^16-1, 
                 ifelse(is.double(db.attr[,2]), min(db.attr[,2]),NA))
db.attr <- db.attr %>%
  na_if(., na.val)
# summary(db.attr)
```

## Depth of interest {#depth}

```{r depth.list}
depth.list.sg <- c("0-5", "5-15", "15-30", "30-60", "60-100", "100-200")
depth.list.gsm <- c("000_005", "005_015", "015_030", "030_060", "060_100", "100_200")
```

*Select a depth slice* by its position in the list, based on the YAML or run-time parameter, and make a full name from the property of interest and the selected depth slice:

```{r depth}
depth.gsm <- depth.list.gsm[params$depth.n]
(voi.depth.name <- paste0(voi.name, depth.gsm))
```


## Area of Interest (AOI) {#aoi}

Specify the lower-right corner from the YAML or rendering parameters:

```{r lrc}
tile.lrc <- c(params$lrc_long, params$lrc_lat) # lower-right corner
```

Compute the upper-left corner:

```{r tile.1}
# Tile size, in integer degrees
size.long <- params$size; size.lat <- params$size
tile.ulc <- c(tile.lrc[1]-size.long, tile.lrc[2]+size.lat) # upper-left corner
m <- matrix(c(tile.ulc[1],tile.lrc[1],  #ulc
              tile.ulc[2], tile.lrc[2]  #lrc
              ),
            nrow=2)
bb.ll <- st_sfc(st_multipoint(m))
st_crs(bb.ll) <- 4326
```

# Import the grid

Information:

```{r}
grid.name <- paste0(base.dir.gsm05.import, "mukeyint_SaS_3as/mukeyint_SaS_3as.tif")
file.info(grid.name)[c("size", "mtime")]
```

Import the grid:

```{r read.grid}
r <- rast(grid.name)
ext(r)
crs(r)
dim(r)
res(r)
res(r)*60*60
```

Resolution in meters:

```{r}
res(r)[1]*(10000/90)*1000  # at the equator
res(r)[1]*(10000/90)*1000*cos((pi/180)*30)
```

Rename the data item to match the database:

```{r name.grid}
names(r) <- "mukeyint"
```

This is 3' resolution in geographic coordinates.

## Clip to the AOI

Clip to the AOI:

```{r clip.grid}
bb.vect <- as.vector(matrix(st_bbox(bb.ll), nrow=2, byrow=T))
r.crop <- terra::crop(r, bb.vect)
dim(r.crop)
length(unique(r.crop))
head(r.crop, rows=8, cols=8)
```

There are `r length(unique(r.crop))` unique map unit IDs in this window. The upper-left corner is shown with the `head` function.

# Link to attribute of interest

Link these values with the grid cell, via the map unit key `mu_key` which is the first column in the reduced data table.
Determine the substitution value column number from the name.

```{r link, fig.cap="Raster map of selected attribute, GSM v0.5"}
m.reclass <- data.frame(from=db.attr[,1],
                    to=db.attr[, match(voi.depth.name, names(db.attr))])
dim(m.reclass)
m.reclass[1:6,]
system.time(
  r.attr <- classify(r.crop, m.reclass)
)
plot(r.attr)
summary(r.attr)
```

# Save tile {#save}

Save this map for further processing, e.g., comparing with SoilGrids250 or other PSM products.

A prefix for directories, to keep AOI results separate.

```{r}
AOI.dir.prefix <- paste0("lat", tile.lrc[2], tile.ulc[2],
                         "_lon", tile.ulc[1], tile.lrc[1])
```

A directory to store the tile:

```{r save}
dest.dir <-  paste0(base.dir.gsm05, AOI.dir.prefix)
if (!dir.exists(dest.dir)) {
   dir.create(dest.dir, recursive = TRUE)
}
```

Save the tile. Note that the file name includes the property name and depth slice. Specify the integer datatype and a GeoTIFF "world" file.

```{r}
f <- terra::writeRaster(r.attr, file=paste0(dest.dir, "/GSM_",
                                              voi.depth.name, ".tif"),
                        overwrite=TRUE, datatype="INT2U", options=c("TFW=YES"),
                        filetype="GTIFF")
print(f)
```
