---
title: "ISSR8 --- Import for PSM comparisons"
author: "D G Rossiter"
date: "`r Sys.Date()`"
params:
   lrc_long: -76
   lrc_lat: 42 
   size: 1
   voi.n: 4
   quantile.n: NA 
   depth.n: 4
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    number_section: yes
    fig_height: 7
    fig_width: 7
    fig_caption: false
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, purl = FALSE, fig.align = 'center')
knitr::opts_chunk$set(cache.extra = R.version.string, comment="")
```

# Introduction

The [Intermediate-scale gridded soil property and interpretation maps from averaged and aggregated SSURGO and STATSGO data](https://casoilresource.lawr.ucdavis.edu/soil-properties/) from the [California Soil Resource Lab](https://casoilresource.lawr.ucdavis.edu) at UC Davis contains "a variety of soil properties throughout the continental United States ... obtained by aggregating USDA-NCSS soil survey data (SSURGO back-filled with STATSGO where SSURGO is not available) within 800m² grid cells." 

This product is also known as [ISSR-800](https://github.com/ncss-tech/ISSR-800).

The grids were aggregated to 800 x 800 m grid cells:

* Area weighted average of components by map unit
* Area weighted average of all map units within each grid cell

To use this script:

1. Ajust the [directory structure](#dirs) to your system

2. [Select a property](#prop) and [select a depth slice](#depth), using the YAML header or by knitting with parameters.

3. [Select an Area of Interest](#aoi), a $1 \times 1^\circ$ tile, using the YAML header or by knitting with parameters.

4. Either compile to HTML or PDF ("knit"), or "Run All" within R Markdown.

5. The processed tile will be in the directory structure, in a [subdirectory named for the AOI](#import).

This GeoTIFF can then be read into R and compared with other PSM products.


# Packages

```{r}
library(sf)     # Simple Features representation of spatial data
library(terra)  # for raster import and display
```

# Directories {#dirs}

```{r}
base.dir.import <- "/Volumes/Pythagoras/ds/DSM_import/"
base.dir.issr8.import <- paste0(base.dir.import, "ISSR8")
base.dir <- "/Volumes/Pythagoras/ds/DSM_export/"
base.dir.issr8 <- paste0(base.dir, "ISSR8")
```

Make sure the import directory exists:

```{r}
if (!dir.exists(base.dir.issr8.import)) {
   dir.create(dbase.dir.issr8.import,recursive = TRUE)
}
```

# Parameters

Parameters for this run:

```{r}
print(paste("lrc_long:", params$lrc_long, "; lrc_lat:", params$lrc_lat, "; size:", params$size))
print(paste("voi.n:", params$voi.n, "; depth.n:", params$depth.n))
```

## Property of interest {#prop}

List of properties, paralleling SoilGrids:

```{r}
# voi.list.sg <- c("clay", "silt", "sand", "phh2o", "cec", "soc", "bdod", "cfvo")
voi.list.issr8 <- c("clay", "silt","sand","ph", "cec", "", "", "ec", "rf", "paws", "sar")
```

ISSR-800 only has SOM stocks and average bulk density, not by depth slice.

Select a property:

```{r}
voi.issr8 <- voi.list.issr8[params$voi.n]
if (voi.issr8=="") { stop(paste0("No ISSR8 property #", params$voi.n)) }
```

## Depth of interest {#depth}

Depth slices, adjusted to GSM standard depths.

```{r}
depth.list.sg <- c("0-5", "5-15", "15-30", "30-60", "60-100", "100-200")
depth.list.issr8 <- c("05", "025", "2550", "3060", "", "")
```

Select a depth slice. Note only 0-5 and 30-60 can be directly compared to GSM-standard PSM maps. The GSM 5-15 is roughly equivalent to ISSR8 0-25; the GSM 15-30 is roughly equivalent to ISSR8 25-50.
The ISSR8 25-50 is roughly equivalent to GSM 30-60, but we ususally have ISSR8 30-60.

```{r}
depth.issr8 <- depth.list.issr8[params$depth.n]
if (depth.issr8=="") { stop(paste0("No ISSR8 depth slice #", params$depth.n)) }
```

## Area of interest {#aoi}

The AOI is a tile using WGS84 geographic coordinates. A $1 \times 1^\circ$ allows comparison with POLARIS, but here other sizes can be specified.

Specify the _lower-right corner_ and _tile size_ from the YAML or rendering parameters:

```{r lrc}
tile.lrc <- c(params$lrc_long, params$lrc_lat) # lower-right corner
tile.size <- params$size                # tile dimensions
```

Compute the upper-right corner $1^\circ$ west and north:

```{r ulc}
tile.ulc <- c(tile.lrc[1]-tile.size, tile.lrc[2]+tile.size) # upper-left corner
```

A prefix for directories, to keep AOI results separate.

```{r dir.prefix}
AOI.dir.prefix <- paste0("lat", tile.lrc[2], tile.ulc[2],
                         "_lon", tile.ulc[1], tile.lrc[1])
```

Set up a directory for the products, separated by AOI:

```{r}
(dest.dir.issr8 <-  file.path(base.dir.issr8,
                       AOI.dir.prefix))
if (!dir.exists(dest.dir.issr8)) {
   dir.create(dest.dir.issr8,recursive = TRUE)
}
```

# Download

ISSR-800 maps can be interactively
[downloaded from here](https://casoilresource.lawr.ucdavis.edu/soil-properties/download.php)
as all-48 State GeoTIFF.

To download programmatically, use the base URL where the maps are stored:

```{r base.url}
base.url <- "http://soilmap2-1.lawr.ucdavis.edu/800m_grids/rasters/"
```

Note the use of `http` rather than `https` to avoid problems with certificates.

ISSR-800 properties have to be provided at the required depth slice. Some were done by NRCS for their own interests (0-5, 0-25, 25-50), and others can be produced by request.

Build the name of the ISSR8 map. Note not all depths are available for all properties, so check this.

```{r}
issr8.map <- paste0(voi.issr8, "_", depth.issr8, ".tif")
```

Download the all-CONUS raster. These are quite large (8 -- 35 Mb) and therefore are read into the import location. Note that once read in, any number of tiles can be extracted from this.


```{r download.conus}
(dest.file <- paste0(base.dir.issr8.import, "/", issr8.map))
if (!file.exists(dest.file)) {
   download.file(
      url = 
         paste0(base.url, issr8.map), 
      destfile = dest.file,
      extra = "--max-redirect=0",
      method = "auto")
} else {
   print("Local copy of file already exists")
}
```


Read into R and plot:

```{r plot.properties.issr8.import, fig.cap="ISSR-800 tile, all USA"}
r.issr8 <- terra::rast(dest.file)
print(r.issr8)
terra::plot(r.issr8, main=paste("ISSR-800;",  main=voi.issr8))
```

Note the CRS is defined, a CONUS Albers Equal Area.


# Crop to AOI

Make a `terra` `SpatExtent` object from the bounding box. Note we need all four corners to fully cover the $1 \times 1^\circ$ tile.

```{r}
m <- matrix(c(tile.ulc[1], tile.lrc[1], tile.lrc[1], tile.ulc[1],  
              tile.ulc[2], tile.ulc[2], tile.lrc[2], tile.lrc[2]),  
            nrow=4)
bb.ll <- st_sfc(st_multipoint(m))
st_crs(bb.ll) <- 4326   # ESPG code for WGS84 long/lat
```

Project the bounding box to the AEA. This is the CRS used by ISSR-800:

```{r}
crs.aea <- "+proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
(bb.aea <- st_transform(bb.ll, crs.aea))
```

Crop to the bounding box

Make a `terra` `SpatExtent` object from the bounding box and use it to crop.

```{r plot.properties.issr8.crop, fig.cap="ISSR-800 tile, cropped tile"}
st_bbox(bb.aea)
bb.vect <- as.vector(matrix(st_bbox(bb.aea), nrow=2, byrow=T))
bb.aea.ext <- ext(bb.vect)
r.issr8.crop <- terra::crop(r.issr8, bb.aea.ext)
plot(r.issr8.crop, main=paste("ISSR-800;",  main=voi.issr8))
print(r.issr8.crop)
summary(r.issr8.crop)
```

We now have a $1 \times 1^\circ$ tile from ISSR800. Note the NA's can be used as a mask -- these are water areas.

# Save tile {#import}

```{r save.tile}
f <- writeRaster(r.issr8.crop, file=paste0(dest.dir.issr8,"/",
                              voi.issr8, "_", depth.issr8, ".tif"),
            overwrite=TRUE, wopt=list(gdal=c("TFW=YES")),
            filetype="GTiff")
```




