---
title: "LandGIS grids  --- Import for PSM comparisons"
author:
  - "D G Rossiter"
  - "d.g.rossiter@cornell.edu"
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
params:
   lrc_long: -76
   lrc_lat: 42 
   size: 1
   voi.n: 4
   quantile.n: NA 
   depth.n: 4
output:
  html_document:
    fig_align: center
    fig_height: 6
    fig_width: 6
    fig_caption: false
    number_section: yes
    theme: spacelab
    df_print: paged
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      purl = FALSE,
                      fig.align = 'center')
knitr::opts_chunk$set(cache.extra = R.version.string, comment="")
```

# Introduction

[LandGIS](https://opengeohub.org/about-landgis) provides various gridded datasets. These are the products of the [EnvirometriX company](http://envirometrix.nl/) of Wageningen (NL).

For their soil maps, _LandGIS_ follows the approach of SoilGrids V1 (ISRIC), with additional sources of point observations, additional covariates, and  usingensemble machine-learning methods.

_LandGIS_ makes 3D "point" predictions at 6 standard soil depths (0, 10 (not 5 or 15), 30, 60, 100 and 200 cm). These do not correspond directly to several other PSM products, e.g., SoilGrids and GSM v0.5, which use GlobalSoilMap standard depth intervals.

To use this script:

1. Adjust the [directory structure](#dirs) to your system

2. [Select a property](#prop)  and [select a depth slice](#depth), using the YAML header or by knitting with parameters.

3. [Select an Area of Interest](#aoi), using the YAML header or by knitting with parameters.

4. Either compile to HTML or PDF ("knit"), or "Run All" within R Markdown.

5. The processed tile will be in the local directory structure, in a [subdirectory named for the AOI](#import).

This GeoTIFF can then be read into R and compared with other PSM products.

# Packages

```{r}
options("rgdal_show_exportToProj4_warnings"="none") 
library(zen4R)          # access Zenodo
library(sf)             # spatial data types 
library(terra)          # raster data, replaces `raster`
```


# Directories {#dirs}

Set a base directory, specific to the local file system. This is where downloaded LandGIS GeoTIFF are located,

```{r}
base.dir <- "/Volumes/Pythagoras/ds/DSM_export/"
base.dir.landgis <- paste0(base.dir, "LandGIS")
base.dir.import <- "/Volumes/Pythagoras/ds/DSM_import/"
base.dir.landgis.import <- paste0(base.dir.import, "LandGIS")
```

# Parameters

Parameters for this run:

```{r}
print(paste("lrc_long:", params$lrc_long, "; lrc_lat:", params$lrc_lat, "; size:", params$size))
print(paste("voi.n:", params$voi.n, "; depth.n:", params$depth.n))
```


## Variable of interest {#prop}

The grids can be downloaded from [Zenodo](https://zenodo.org/), either manually or (as here) programmatically.

For example here is the Zenodo page for [silt concentration](https://zenodo.org/record/2525676). This shows links for all the depths and quantiles.

Search for these coverages as grids (GeoTIFF)
[here in Zenodo](https://zenodo.org/search?page=1&size=20&q=landgis%20soil&file_type=tif).
Files are named as explained [here](https://github.com/Envirometrix/LandGISmaps#soil-properties-and-classes) and on the Zenodo page for each property:

> "For soil variable names we use consistently the National Cooperative Soil Characterization Database column names and codes. For example:
<br>
`sol_bulkdens.fineearth_usda.4a1h_m_250m_b30..30cm_1950..2017_v0.2.tif`
<br>
refers to the `db_od` column in the database and `4a1h` laboratory method (bulk density oven-dry), as described in the [Laboratory Methods Manual (SSIR 42) - NRCS - USDA](https://www.nrcs.usda.gov/Internet/FSE_DOCUMENTS/nrcseprd1026806.pdf)
and the [Soil Survey Field and Laboratory Methods Manual - NRCS - USDA](https://www.nrcs.usda.gov/Internet/FSE_DOCUMENTS/stelprdb1244466.pdf)
<br>
Continuous 3D soil properties are predicted at 6 standard depths: 0, 10, 30, 60, 100 and 200 cm.
Standard prediction errors (either as the prediction variance of prediction confidence limits) are provided for each soil property / depth."

Examples:

`sol_sand.wfraction_usda.3a1a1a_m_250m_b60..60cm_1950..2017_v0.2`: sand, using the `usda.3a1a1a` method, the `m`ean, `250m` resolution, 60 cm depth, data from 1950 to 2017, version 0.2.

Replacing `m`ean with `md`, i.e., mean deviation.

According to the data store six properties are predicted by LandGIS:

This part of the name specifies the property:

* clay.wfraction_usda.3a1a1a
* silt.wfraction_usda.3a1a1a
* sand.wfraction_usda.3a1a1a
* ph.h2o_usda.4c1a2a
* organic.carbon_usda.6a1c
* bulkdens.fineearth_usda.4a1h
* coarsefrag.vfraction_usda.3b1
* watercontent.33kPa_usda.4b1c
* texture.class_usda.tt

The last two have no SoilGrids equivalent.

Zenodo IDs:

* clay: stable 1476885; latest 2525663
* silt: stable 2525675; latest 2525676
* sand: stable 1476851; latest 2525662
* ph.h2o: stable 1475459; latest 252664
* organic.carbon: stable 1475458; latest 2525553
* bulkdens.fineearth: stable 1475971; latest 2525665
* coarsefrag.vfraction: stable 2525681; latest 2525682

Use the stable Zenodo codes to get the most recent version.

```{r}
voi.list.landgis <- c("clay.wfraction_usda.3a1a1a",
                      "silt.wfraction_usda.3a1a1a",
                      "sand.wfraction_usda.3a1a1a",
                      "ph.h2o_usda.4c1a2a",
                      "organic.carbon_usda.6a1c",
                      "bulkdens.fineearth_usda.4a1h",
                      "coarsefrag.vfraction_usda_3b1")
voi.list.landgis.code <- c(1476854, 2525675, 1476851,
                           1475459, 1475457, 1475970,
                           2525681)
```


Select the position in these lists, from the YAML or dynamic rendering:

```{r}
voi.landgis <- voi.list.landgis[params$voi.n]
voi.landgis.code <- voi.list.landgis.code[params$voi.n]
```

## Depth slice {#depth}

LandGIS are point predictions within the profile, conceptually the value of an infinitely thin slice. 
To compare with GlobalSoilMap standards they must be averaged. 
There is no direct correspondence for the 0-5, 5-15, 15-30 slices.
So we do the following:

* 0-5: weighted average of 0 and 10, giving 0 2x the weight of 10
* 5-15: use the 10 cm point prediction -- in fact SoilGrids250 does this
* 15-30: weighted average of 10 and 30, giving 30 2x the weight of 10

```{r}
depth.list.sg <- c("0-5", "5-15", "15-30", "30-60", "60-100", "100-200")
depth.list.landgis <- c("0", "10", "30", "60", "100", "200")
```

Select a slice:

```{r set.depth.slices}
depth.sg <- depth.list.sg[params$depth.n]
if (params$depth.n == 1) {
  top <- 1; bottom <- 2; weights <- c(2/3, 1/3)  
} else if (params$depth.n == 2) { # both the same
  top <- bottom <- 2; weights <- c(1/2, 1/2)
} else if (params$depth.n == 3) {
  top <- 2; bottom <-3; weights <- c(1/3, 2/3) 
} else if (params$depth.n == 4) {
  top <- 3; bottom <- 4; weights <- c(1/2, 1/2)
} else if (params$depth.n == 5) {
  top <- 4; bottom <- 5; weights <- c(1/2, 1/2)
} else if (params$depth.n == 6) {
  top <- 5; bottom <- 6; weights <- c(1/2, 1/2)
}
```

Build the names of the LandGIS files corresponding to the slice:

```{r slice.file.names}
depth.landgis.top <- depth.list.landgis[top]
depth.landgis.bottom <- depth.list.landgis[bottom]
depth.landgis.suffix.top <- paste0("b", depth.landgis.top, "..", depth.list.landgis[top], "cm")
depth.landgis.suffix.bottom <- paste0("b", depth.landgis.bottom, "..", depth.list.landgis[bottom], "cm")
```


## Area of Interest (AOI) {#aoi}

Specify the lower-right corner from the YAML or rendering parameters:

```{r lrc}
tile.lrc <- c(params$lrc_long, params$lrc_lat) # lower-right corner
tile.size <- params$size
```

Compute the upper-left corner:

```{r ulc}
tile.ulc <- c(tile.lrc[1]-tile.size, tile.lrc[2]+tile.size) # upper-left corner
```


# Download LandGIS layers 

Import via Zenodo. Note that this imports *all* the layers for the property, both the mean and the mean deviation, even though we only specify the mean (in the filename) and one depth. So if the property has not been downloaded, this is quite slow. 

Names look like `sol_ph.h2o_usda.4c1a2a_m_250m_b0..0cm_1950..2017_v0.2`

This is very slow, so check if the file has already been downloaded.

*Note* if this download fails, you will have to download manually into the import directory.

```{r zenodo.new}
(filename.in.top <- paste0(base.dir.landgis.import, 
                          "/sol_", voi.landgis, 
                          "_m_250m_", depth.landgis.suffix.top,
                          "_1950..2017_v0.2.tif"))
if (!file.exists(filename.in.top)) {
  system.time(
  download_zenodo(paste0("10.5281/zenodo.",voi.landgis.code),
                  path=base.dir.landgis.import,
                  logger="INFO")
  ) } else print("File previously downloaded")
(filename.in.bottom <- paste0(base.dir.landgis.import, 
                          "/sol_", voi.landgis, 
                          "_m_250m_", depth.landgis.suffix.bottom,
                          "_1950..2017_v0.2.tif"))
if (!file.exists(filename.in.bottom)) {
  system.time(
  download_zenodo(paste0("10.5281/zenodo.",voi.landgis.code),
                  path=base.dir.landgis.import,
                  logger="INFO")
  ) } else print("File previously downloaded")
```

MD5 Checksum to compare with that listed on the Zenodo file list:

```{r}
tools::md5sum(filename.in.top)
tools::md5sum(filename.in.bottom)
```

e.g., for mean pH at 0 cm this is `c4315549f2abdcacd8f58c775de52088`.

# Import LandGIS layer to R

```{r in.slice1}
r.top <- terra::rast(filename.in.top)
r.bottom <- terra::rast(filename.in.bottom)
print(r.top)
```

# Crop to the bounding box {#crop}

Make a `terra` `SpatExtent` object from the bounding box and use it to crop.

Bounding box in geographic coordinates in WGS84:

```{r}
m <- matrix(c(tile.ulc[1],tile.lrc[1],  #ulc
              tile.ulc[2], tile.lrc[2]),  #lrc
            nrow=2)
bb.ll <- st_sfc(st_multipoint(m))
st_crs(bb.ll) <- 4326
bb.vect <- as.vector(matrix(st_bbox(bb.ll), nrow=2, byrow=T))
bb.ll.ext <- ext(bb.vect)
```

Crop to this extent:

```{r crop}
r.top.crop <- terra::crop(r.top, bb.ll.ext)
r.bottom.crop <- terra::crop(r.bottom, bb.ll.ext)
```

# Compute depth slice

Average according to the weighting:

```{r average}
r.landgis.crop <- r.top.crop*weights[1] + r.bottom.crop*weights[2]
```

Show the cropped layer:

```{r plot.properties, fig.cap="Cropped LandGIS tile"}
print(r.landgis.crop)
summary(r.landgis.crop)
terra::plot(r.landgis.crop,
            main=paste0("LandGIS, ", voi.landgis, ", depth ", depth.sg))
```

# Save the tile {#import}

Save to a local directory, using the AOIas subdirectories, and the variable of interest and depth slice in the file name.


A prefix for directories, to keep AOI results separate.

```{r}
AOI.dir.prefix <- paste0("lat", tile.lrc[2], tile.ulc[2],
                         "_lon", tile.ulc[1], tile.lrc[1])
```

Destination directory:

```{r}
(dest.dir.landgis <-  paste0(base.dir.landgis, "/", AOI.dir.prefix))
if (!dir.exists(dest.dir.landgis)) {
  dir.create(dest.dir.landgis, recursive = TRUE)}
```
    
```{r}
(file.name <- paste0(voi.landgis, "_", depth.sg))
f <- writeRaster(r.landgis.crop, file=paste0(dest.dir.landgis,"/",
                                             file.name, ".tif"),
                 overwrite=TRUE, wopt=list(gdal=c("TFW=YES")),
                 filetype="GTiff")
```



