---
title: "POLARIS --- Import for PSM comparisons"
author:
  - "D G Rossiter"
  - "d.g.rossiter@cornell.edu"
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
params:
   lrc_long: -76
   lrc_lat: 42 
   size: 1
   voi.n: 4
   depth.n: 4
   quantile.n: 3 
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    number_section: yes
    fig_height: 7
    fig_width: 7
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, purl=FALSE, fig.align = 'center')
knitr::opts_chunk$set(cache.extra = R.version.string)
```

# Introduction

_POLARIS_ is the result of harmonizing diverse SSURGO and STATSGO polygon data with the DSMART algorithm to produce a raster soil series map (30 m resolution) and then extracting property information from Official Series Descriptions.

It is explained in: Chaney, N. W., Minasny, B., Herman, J. D., Nauman, T. W., Brungard, C. W., Morgan, C. L. S., McBratney, A. B., Wood, E. F., & Yimam, Y. (2019). POLARIS soil properties: 30-m probabilistic maps of soil properties over the contiguous United States. Water Resources Research, 55(4), 2916–2938. Scopus. https://doi.org/10.1029/2018WR022797 

POLARIS is organized as $1 \times 1^\circ$ tiles.

To use this script:

1. Ajust the [directory structure](#dirs) to your system

2. [Select a property](#voi) and [select a depth slice](#depth).

3. [Select an Area of Interest](#aoi).

4. Either compile to HTML or PDF ("knit"), or "Run All" within R Markdown.

5. The processed tile will be in the directory structure, in a [subdirectory named for the AOI](#import).

This GeoTIFF can then be read into R and compared with other PSM products.

# Directories {#dirs}

Set base directories, specific to the local file system. 

1. `base.dir.import`: This is where downloaded large GeoTIFF are located. Because of their size they may be on a separate file system, e.g., removable or networked drive.

2. `base.dir`: This is where the processed POLARIS maps are stored.


```{r}
base.dir <- "/Users/rossiter/ds/"
base.dir.polaris <- paste0(base.dir, "POLARIS/")
base.dir.import <- "/Volumes/Pythagoras/ds/"
base.dir.polaris.import <- paste0(base.dir.import, "POLARIS/")
```

These are the base of destination directories built [below](#dest)

# Packages

```{r}
library(terra)  # for raster import and display
```


# Variable and quantile of interest {#voi}

Define the variables for the soil property and layer of interest.

`p5` - 5% quantile from the machine-learning methods
`p50` - median of the distribution
`p95` - 95% quantile
`mean` - mean of the distribution;

```{r}
quantile.list.polaris <- c("p5", "p50", "p95", "mean")
```

The list of properties and their units of measure is [here](http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/Readme).

Relevant here are:

* clay - clay percentage, %
* silt - silt percentage, %
* sand - sand percentage, %
* ph - soil pH in H2O, N/A
* om - organic matter, log10(%)
* bd - bulk density, g/cm3

There are also soil hydrological parameters we do not process.

```{r}
voi.list.polaris <- c("clay", "silt", "sand", "ph", "", "om", "bd", "") 
```

Set the property and quantile from the YAML or rendering parameters:

```{r}
voi.polaris <- voi.list.polaris[params$depth.n]
quantile.polaris <- quantile.list.polaris[params$quantile.n]
```


# Depth slice {#depth}

Depth slices, in cm:

```{r}
depth.list.polaris <- c("0_5", "5_15", "15_30", "30_60", "60_100", "100_200")
```

Set the depth from the YAML or rendering parameters.
Then build a full layer name with the property, depth, and quantile:

```{r}
depth.polaris <- depth.list.polaris[params$depth.n]
(voi_layer <- paste(voi.polaris, depth.polaris, quantile.polaris, sep="_")) # layer of interest 
```

# Area of Interest (AOI) {#aoi}

POLARIS data is served in $1 \times 1^\circ$ tiles using WGS84 geographic coordinates. Specify the _lower-right corner_, then compute the upper-right corner $1^\circ$ west and north.

Specify the lower-right corner from the YAML or rendering parameters:

```{r lrc}
tile.lrc <- c(params$lrc_long, params$lrc_lat) # lower-right corner
```

Compute the upper-left corner:

```{r ulc}
tile.ulc <- c(tile.lrc[1]-1, tile.lrc[2]+1) # upper-left corner
```

A prefix for directories, to keep AOI results separate.

```{r dir.prefix}
AOI.dir.prefix <- paste0("lat", tile.lrc[2], tile.ulc[2],
                         "_lon", tile.ulc[1], tile.lrc[1])
```

# Import the POLARIS tile {#import}

POLARIS tiles as GeoTIFFs can be downloaded [here](http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0).
These are organized as hierarchical directory: property, depth, quantile.
The file name contains the tile bounding box, e.g. `lat4142_lon-78-77.tif`.

Download the file if we don't already have it.


Second, make sure the directory exists, creating it if necessary.

```{r}
(dest.dir.polaris.import <-  paste0(base.dir.polaris.import, 
                            AOI.dir.prefix, "/",
                            voi.polaris, "/", 
                            quantile.polaris, "/", 
                            depth.polaris))
if (!dir.exists(dest.dir.polaris.import)) {
   dir.create(dest.dir.polaris.import, recursive = TRUE)
}
```

Finally, download the tile.

```{r}
polaris.tile <- paste0("lat", tile.lrc[2], tile.ulc[2],
                        "_lon", tile.ulc[1], tile.lrc[1],
                        ".tif")
(dest.file <- paste0(dest.dir.polaris.import, "/", polaris.tile))
if (!file.exists(dest.file)) {
   download.file(
      url = 
         paste0("http://hydrology.cee.duke.edu/POLARIS/PROPERTIES/v1.0/",
                voi.polaris, "/", quantile.polaris, "/", depth.polaris, "/",
                polaris.tile), 
      destfile = dest.file,
      method = "auto")
} else {
   print("Local copy of file already exists")
}
```

These are quite large (8 -- 35 Mb) and therefore are read into the import location.
They can be deleted after they have been read into R.

# Check

Import to R and print/plot, to check:

```{r}
r <- terra::rast(dest.file)
print(r)
plot(r)
```

