---
title: "Comparing PSM products at local resolution"
author:
- D G Rossiter
- d.g.rossiter@cornell.edu
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
params:
   lrc_long: -86
   lrc_lat: 38
   size: 1
   voi.n: 6
   quantile.n: NA
   depth.n: 1
   test.tile.size: 0.15  # degrees
   test.tile.x.offset: 0.61  # west from right edge
   test.tile.y.offset: 0.45  # north from bottom edge
output:
  html_document:
    fig_align: center
    fig_height: 6
    fig_width: 6
    number_section: yes
    theme: spacelab
    df_print: paged
    code_folding: hide
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      purl=FALSE,
                      fig.align = 'center',
                      fig.path = './figs/compare_30m/')
knitr::opts_chunk$set(cache.extra = R.version.string)
```

# Introduction

This script compares PSM products at 30 m grid resolution, which is used by POLARIS, and is a common resolution for local products. A grid cell is thus about 0.09 ha, sufficient for field-specific assessment even for small fields.

These two are  compared:

* [gSSURGO](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/geo/?cid=nrcseprd1464625);
* POLARIS Soil Properties;

And this is added as a representative global product, downscaled to the local resolution:

* [SoiLGrids250](https://www.isric.org/explore/soilgrids)  from ISRIC further abbreviated as _SG2_.

The PSM products must have been previously imported and restricted to the same area of interest (AOI), a $1 \times 1^\circ$, to the locations indicated in the directory list. These can be somewhat larger, in this script the CRS are made compatible (WGS84 geographic) and cropped to exactly a 1 degree tile.

We use gSSURGO as the reference map.

This script must follow the import of the various products; these are in directory `../scripts_importmaps`.

To use this script:

1. Ajust the [directory structure](#dirs) to your system.

2. [Select a property](#voi) and [select a depth slice](#depth), using the YAML header or by knitting with parameters..

3. [Select an Area of Interest](#aoi), using the YAML header or by knitting with parameters.

4. [Set the sub-tile](#sub_aoi), if you want to compare a smaller area than a full tile; this makes sense since we want to see details at this resolution.

The above can be adjusted in the YAML header; these include the default parameters and look like:

```
---
   params:
   lrc_long: -76
   lrc_lat: 42 
   size: 1
   voi.n: 4
   quantile.n: NA 
   depth.n: 4
   test.tile.size: 0.15  # degrees
   test.tile.x.offset: 0.32  # west from right edge
   test.tile.y.offset: 0.68  # north from bottom edge
---
```

These can also be specified with the `params` argument in a call to `rmarkdown::render`.

5. Either compile to HTML or PDF ("knit"), or "Run All" within R Markdown.

Results:

1. Generated figures will be in directory `./figs/compare_local/`. These will be specific to the AOI, property and depth slice. So to save them from over-writing by other runs, move them to another directory.

2. Generated tables in \LaTeX format will be in directory `../LaTeX_tables`. The table names include AOI, property and depth slice and so are not over-written.

3. Generated harmonized maps will be [saved](#save) to directory `Compare_30m` under the base directory, in a subdirectory named for the AOI. The file name includes the PSM method, property and depth slice.


# Setup

This is a simplified version of the code in `Compare_regional.Rmd`.

```{r figure.setup}
n.products <- 3
n.figs.row <- 1
n.figs.col <- 3
map.fig.width <- n.figs.col*5
map.fig.height <- n.figs.row*5
n.figs.row.diff <- 1
n.figs.col.diff <- 2
map.fig.width.diff <- n.figs.col.diff*5
map.fig.height.diff <- n.figs.row.diff*5
```


## Packages

```{r}
library(rgdal)      # R interface to GDAL
library(terra)      # for raster maps
library(sf)         # Simple Features spatial data
# library(gridExtra)  # arrange multiple plots
library(knitr)      # for fancy tables
library(xtable)     # (same)
```

## Base directory paths {#dirs}

Set base directories, specific to the local file system. 

1. `base.dir`: This is the location of the PSM tiles that have been cropped to an AOI by an import script.

```{r base.dir}
base.dir <- "/Volumes/Pythagoras/ds/DSM_export"
base.dir.gssurgo <- paste0(base.dir, "gSSURGO")
base.dir.polaris <- paste0(base.dir, "POLARIS")
base.dir.sg <- paste0(base.dir, "SoilGrids250")
```

2. `base.dir.import`: This is where downloaded large GeoTIFF are located. Because of their size they may be on a separate file system, e.g., removable or networked drive. Files may have been downloaded here by an import script, or by direct download from the data provider.

In this script this location is only used for POLARIS, since these large files (15-30Mb) are imported directly as $1 \times 1^\circ$ tiles by script `./scripts_importmaps/POLARIS_import.Rmd`.

```{r base.dir.import}
base.dir.import <- "/Volumes/Pythagoras/ds/"
base.dir.polaris.import <- paste0(base.dir.import, "POLARIS")
```

3. `base.dir.export`: This is where generated harmonized maps will be saved for further processing, e.g., comparing patterns.

These are quite large files, about 50Mb for a 1-degree tile, so store on the external storage.

```{r base.dir.export}
base.dir.export <- paste0(base.dir.import, "Compare_PSM_local")
```

# Parameters

Parameters for this run:

```{r}
print(paste("lrc_long:", params$lrc_long, "; lrc_lat:", params$lrc_lat, "; size:", params$size))
print(paste("voi.n:", params$voi.n, "; depth.n:", params$depth.n))
print(paste("test.tile.size:", params$test.tile.size, 
            "test.tile.x.offset:", params$test.tile.x.offset,
            "test.tile.y.offset:", params$test.tile.y.offset))
```

## Property of interest {#voi}

Property names in various systems. Note that all except GSM v0.5 and SoilGrids250 are missing one or more properties. Such products will be omitted from comparisons for those properties.

Properties that be compared are: clay, silt, sand weight concentrations; pH in 1:1 water; CEC; SOC; bulk density of the fine earth; coarse fragment volume proportion. See the import script for each product for a link to a description of the properties and their units of measure.

```{r}
voi.list.gssurgo <- c("claytotal_r", "silttotal_r", "sandtotal_r",
                  "ph1to1h2o_r", "cec7_r", "om_r",   # note SOM not SOC
                  "dbthirdbar_r", "sieveno10_r") # passing 2.0 mm sieve, complement is coarse fragments
voi.list.sg <- c("clay", "silt", "sand", "phh2o", "cec", "soc", "bdod", "cfvo")
voi.list.polaris <- c("clay", "silt", "sand", "ph", "", "om", "bd", "") 
```

Select the position in these lists

```{r}
voi.n <- params$voi.n   # variable of interest, SoilGrids name
voi.gssurgo <- voi.list.gssurgo[voi.n]
voi.sg <- voi.list.sg[voi.n]
voi.polaris <- voi.list.polaris[voi.n]
```

## Depth of interest {#depth}

Depth slices:

```{r}
depth.list.gssurgo <- c("05", "515", "1530", "3060", "60100", "100200")
depth.list.sg <- c("0-5", "5-15", "15-30", "30-60", "60-100", "100-200")
depth.list.polaris <- gsub("-", "_", depth.list.sg)
```

Select the depth slice:

```{r}
depth <- params$depth.n
```


## Area of Interest (AOI) {#aoi}

We use a $1 \times 1^\circ$ tile, because that is how POLARIS data is served.

Specify the _lower-right corner_ and _tile size_ from the YAML or rendering parameters:

```{r lrc}
tile.lrc <- c(params$lrc_long, params$lrc_lat) # lower-right corner
tile.size <- params$size                # tile dimensions
```

Compute the upper-right corner $1^\circ$ west and north:

```{r ulc}
tile.ulc <- c(tile.lrc[1]-tile.size, tile.lrc[2]+tile.size) # upper-left corner
```

A prefix for directories and file names, to keep AOI results separate.

```{r aoi.dir.prefix}
AOI.dir.prefix <- paste0("lat", tile.lrc[2], tile.ulc[2],
                         "_lon", tile.ulc[1], tile.lrc[1])
```

Change the location of figures generated by this script: put in subdirectories by area, 

```{r adjust.fig.path}
knitr::opts_chunk$set(fig.path = paste0(knitr::opts_chunk$get("fig.path"), 
                                        AOI.dir.prefix, "/",
                                        voi.sg, "_", depth.list.sg[depth], "_"))
```

Bounding box:

```{r bbox.4326}
m <- matrix(c(tile.ulc[1],tile.lrc[1],  #ulc
              tile.ulc[2], tile.lrc[2]), nrow=2) #lrc
bb.ll <- st_sfc(st_multipoint(m))
st_crs(bb.ll) <- 4326   # ESPG code for WGS84 long/lat
```

Project the bounding box to Goode Interrupted Homolosine (IGH) used by SoilGrids:

```{r bbox.igh}
# convert to Homolosine. Note epsg=152160 is not in PROJ4 database
crs.igh <- '+proj=igh +lat_0=0 +lon_0=0 +datum=WGS84 +units=m +no_defs'
(bb.igh <- st_transform(bb.ll, crs.igh))
(bb.igh <- st_coordinates(bb.igh)[,1:2])
(bb <- as.vector(t(bb.igh)))
```

Project the bounding box to the CONUS AEA. This is the CRS used by gSSURGO:

```{r bbox.aea}
crs.aea <- "+proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
(bb.aea <- st_transform(bb.ll, crs.aea))
```


## Test area {#sub_aoi}

Because the full tile is so large, here we can choose to test on a smaller area, e.g., a quarter-tile 0.25 x 0.25 degrees.

Set the size and offset from the lower-right corner. The test area will be to the north and west of this.

```{r}
tile.lrc <- c(params$lrc_long, params$lrc_lat) # lower-right corner
tile.size <- params$size     
test.tile.size <- params$test.tile.size  # degrees
test.tile.x.offset <- params$test.tile.x.offset  # west from right edge
test.tile.y.offset <- params$test.tile.y.offset  # north from bottom edge
## CNY
# test.tile.size <- 0.15  # degrees
# test.tile.x.offset <- 0.30  # west from right edge
# test.tile.y.offset <- 0.45  # north from bottom edge
## NC
# test.tile.size <- 0.15  # degrees
# test.tile.x.offset <- 0.32  # west from right edge
# test.tile.y.offset <- 0.68  # north from bottom edge
## IN
# test.tile.size <- 0.15  # degrees
# test.tile.x.offset <- 0.61  # west from right edge
# test.tile.y.offset <- 0.44  # north from bottom edge
```


# Products to compare

Load the tiles in their native CRS, as processed in the import scripts.

## gSSURGO

This is the reference product.

```{r get.tiles.gssurgo}
src.dir <-  paste0(base.dir.gssurgo ,"/", 
                   AOI.dir.prefix)
(voi.depth.name <- paste0(voi.gssurgo, "_", depth.list.gssurgo[depth]))
(file.name <- paste0(src.dir, "/", voi.depth.name, '.tif'))
if (file.exists(file.name)) {
  r.gssurgo <- terra::rast(file.name)
  names(r.gssurgo) <- "gssurgo"
  print(r.gssurgo)
} else { stop("No gSSURGO tile, stopping") }
```


## SoilGrids250

This is the globally-consistent product.

```{r get.tiles.sg}
# SoilGrids250 -- only the mean prediction in this script
# Use the EPSG:4326 version
src.dir <-  paste0(base.dir.sg ,"/", 
                   AOI.dir.prefix, "/", 
                   voi.sg, "/mean/",
                   depth.list.sg[depth], "cm")
(voi.depth.name <- paste0(voi.sg, "_", depth.list.sg[depth], "cm_mean_4326"))
(file.name <- paste0(src.dir, "/", voi.depth.name, '.tif'))
if (file.exists(file.name)) {
  r.sg <- terra::rast(file.name)
  names(r.sg) <- "SoilGrids250"
  print(r.sg)
} else { stop("No SoilGrids250 tile, stopping") }
```


## POLARIS

```{r get.tiles.polaris}
# POLARIS -- only the mean prediction in this script
(file.name <- paste0(base.dir.polaris.import, "/",
                     AOI.dir.prefix, "/",
                     voi.list.polaris[params$voi.n], "/mean/",
                     depth.list.polaris[depth], "/",
                     AOI.dir.prefix, ".tif"))
if (file.exists(file.name)) {
  r.p <- terra::rast(file.name)
  names(r.p) <- "polaris"
  print(r.p)
}  else { stop("No POLARIS tile, stopping") }
```


# Make maps compatible

## Make the units compatible

Depending on the property, data in some coverages need to be converted to the units used in SoilGrids250; we choose this as the base units; note these are integers. Here are the units:

 
```{r show.conversions}
df <- data.frame(property=voi.list.sg, 
                 #"clay"  "silt"  "sand"  "phh2o" "cec"   "soc"   "bdod"  "cfvo" 
                 sg=c("%%","%%","%%","pHx10","mmol(c)/kg","dg/kg","cg/cm3", "cm3/dm3"),  #SG
                 ## metadata for gSSURGO:
                 ## https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/geo/?cid=nrcs142p2_053631
                 gssurgo=c("%","%","%","pH","meq/Cg","1.724 * %","g/cm3", "100-%"),  # gSSURGO
                  # sieveno10_r is complement of cfvo; som not soc
                 p=c("%","%", "%","pH","", "log10(1.724 * %)", "g/cm3", ""))  # POLARIS
 knitr::kable(
  df, caption = 'Properties and units of measure',
  col.names=c("Property", "SoilGrids", "gSSURGO",  "POLARIS"),
  booktabs = TRUE)
```  

Make a matrix with the conversions to SoilGrids250 units. These factors *multiply* the source, to match SoilGrids250. `NA` values indicate that the property is not included in the source, or that its conversion can not be handled by multiplication.

Some conversions are given [here](https://www.isric.org/explore/soilgrids/faq-soilgrids#What_do_the_filename_codes_mean).

```{r make.conversion.matrix}
som.to.soc <- 1/1.724138 # this was used in the lab, I know it has been heavily criticized
conversions <- data.frame(property=voi.list.sg, 
                 # sg=c("%%","%%","%%","pHx10","mmol(c)/kg","dg/kg","cg/cm3", "cm3/dm3"), #SG
                 gssurgo=c(10, 10, 10, 10 , 10, 1000*som.to.soc, 100, NA),  # gSSURGO
                 p=c(10,  10, 10,  10, NA, NA, 100, NA)  # POLARIS -- SOM is special case
                 )
conversions <- format(conversions, digits=4, nsmall=0, scientific=FALSE)
knitr::kable(
  conversions, caption = 'Conversion factors, multiply by these to match SoilGrids250',
  col.names=c("Property", "gSSURGO", "POLARIS"),
  booktabs = TRUE,
  align = "r")
```

Convert units as necessary.  Use SoilGrids as the default.

SOC for POLARIS is a special case, because of the log10-scale, and because it is SOM, not SOC. Use the conventional conversion factor 0.58 = 1/1.724138.

```{r polaris.soc}
if (exists("r.p") && (voi.sg=="soc")) {
    r.p <- ((10^r.p)*som.to.soc*1000) 
}
```

Coarse fragments for gSSURGO is another special case:

```{r gssurgo.cfvo}
if (voi.sg == "cfvo") {
  r.gssurgo <- (100 - r.ssurgo)*0.1
}
```


All other conversions:


```{r convert}
# this property's factors
(factors <- conversions[match(voi.sg, conversions$property),])

# gSSURGO
fact <- as.numeric(factors["gssurgo"])
if (!is.na(fact) && (fact != 1)) { r.gssurgo <- r.gssurgo*fact }

fact <- as.numeric(factors["p"])
if (!is.na(fact) && (fact != 1)) { r.p <- r.p*fact }
```



## Match resolution and CRS

SoilGrids250 and POLARIS are in EPSG:4326 (WGS84 long/lat), at different grid resolutions.

```{r crs.show.1}
rgdal::showP4(crs(r.sg))
data.frame(sg=res(r.sg)[1], 
           polaris=res(r.p)[1])
```

gSSURGO is in a CONUS Albers Equal Area on the GRS80 ellipsoid, the basis of the NAD83 datum, and centred with WGS84, so there is effectively no difference in datums with the other products.

```{r crs.show.2}
rgdal::showP4(crs(r.gssurgo))
```

Resample SoilGrids nominal 250m into POLARIS WGS84 30m nominal pixels, using cubic interpolation -- this will only have effect near the edges of the large pixel being downscaled.

```{r crs.resample}
r.sg.p <- terra::resample(r.sg, r.p, method="cubic")
# plot(r.sg.p)
```

For gSSURGO the CRS must be changed, not just the resolution.
Since the resolutions are similar (30 m vs. 1 arc-second), use nearest-neighbour to avoid diffuse values along what had been polygon boundaries in the original map.The value range will not change.

This avoids spurious very thin polygons, if the raster has to be polygonized, e.g., as class maps.

```{r crs.project}
r.gssurgo.p <- terra::project(r.gssurgo, r.p, method="near") 
# plot(r.gssurgo.p)
```


## Make all maps cover the same area


Make a true 1 degree tile raster and use it to mask the other coverages, to get consistent coverage.
SoilGrids is larger and so are the two originally in AEA projection.

To be used for cropping it must be a `Spatial` class from which a `terra::SpatExtent` can be extracted.

```{r make.polygon}
m <- matrix(c(tile.ulc[1],tile.ulc[2],  #ulc
              tile.lrc[1],tile.ulc[2],  #urc
              tile.lrc[1],tile.lrc[2],  #lrc
              tile.ulc[1],tile.lrc[2],  #llc
              tile.ulc[1], tile.ulc[2]), nrow=5, byrow = TRUE) # ulc again, closes polygon
bb.poly <- st_sfc(st_linestring(m))
st_crs(bb.poly) <- 4326 
bb.poly <- st_polygonize(bb.poly)
bb.poly <- as_Spatial(bb.poly)
ext(bb.poly)
```

```{r}
r.gssurgo.p <- crop(r.gssurgo.p, bb.poly)
r.p <- crop(r.p, bb.poly)
r.sg.p <- crop(r.sg.p, bb.poly)
```

POLARIS predicts in the lakes, gNATSGO predict in some built-up areas. Mask these out with SoilGrids250, which does not.

```{r mask.with.polaris}
r.gssurgo.p <- mask(r.gssurgo.p, r.sg.p)
r.p <- mask(r.p, r.sg.p)
```

There are some areas not surveyed by gSSURGO, e.g., Indian reservations. Also their unsurveyed urban areas may differ. So mask the other products with gSSURGO.

```{r mask.with.gssurgo}
r.p <- mask(r.p, r.gssurgo.p)
r.sg.p <- mask(r.sg.p, r.gssurgo.p)
```




## Crop to test area


```{r crop.test.area}
(tmp <- as.vector(ext(r.sg.p)))
tmp["xmax"] <- tmp["xmax"] - test.tile.x.offset
tmp["xmin"] <- tmp["xmax"] - test.tile.size
tmp["ymin"] <- tmp["ymin"] + test.tile.y.offset
tmp["ymax"] <- tmp["ymin"] + test.tile.size
ext(tmp)
r.gssurgo.p <- crop(r.gssurgo.p, ext(tmp))
r.sg.p <- crop(r.sg.p, ext(tmp))
r.p <- crop(r.p, ext(tmp))
```


# Compare

## Compute common range for all products

```{r compare.zlim}
zlim <- c(min(values(r.sg)*10, na.rm = TRUE),
          max(values(r.sg)*10, na.rm = TRUE))/10
zlim <- c(floor(min(zlim[1]*10, values(r.gssurgo.p)*10, na.rm=TRUE)),
          ceiling(max(zlim[2]*10, values(r.gssurgo.p)*10, na.rm=TRUE)))/10
zlim <- c(floor(min(zlim[1]*10, values(r.p)*10, na.rm=TRUE)),
          ceiling(max(zlim[2]*10, values(r.p)*10, na.rm=TRUE)))/10
print(zlim)
```

## Histograms
    
The property and depth will be given in the caption, if this is used in a publication.

First, compute the maximum density, to have a common y-axis:

```{r hist.densities}
max.dens <- function(r.map) {  # argument: the raster map
  h <- hist(r.map,  breaks=24, plot = FALSE)
  max(h$counts/(diff(h$breaks[1:2]))/sum(h$counts))
}
yl <- c(0, max(max.dens(r.gssurgo.p), max.dens(r.sg.p), max.dens(r.p)))
```

Use this y-axis to show all the histograms together:

```{r hist.sg.props, fig.width=map.fig.width, fig.height=map.fig.height}
par(mfrow=c(n.figs.row, n.figs.col))
hist(r.gssurgo.p, breaks=24, main="gSSURGO",
     xlim=zlim, xlab="", freq = FALSE, ylim=yl)
hist(r.p, breaks=24, main="POLARIS",
     xlim=zlim, xlab="", freq = FALSE, ylim=yl)
hist(r.sg, breaks=24, main="SoilGrids25O",
     xlim=zlim, xlab="", freq = FALSE, ylim=yl)
par(mfrow=c(1,1))
```

## Maps


```{r map.sg.props, fig.width=map.fig.width, fig.height=map.fig.height}
par(mfrow=c(n.figs.row, n.figs.col))
terra::plot(r.gssurgo.p, main="gSSURGO", range=zlim)
terra::plot(r.p, main="POLARIS", range=zlim)
terra::plot(r.sg.p, main="SoilGrids", range=zlim)
par(mfrow=c(1,1))
```

## Correlations

Pairwise Pearson correlations. Note the distributions are fairly symmetric/quasi-normal so Pearson's correlations are valid.

```{r pairwise}
v.all <- data.frame(gssurgo=values(r.gssurgo.p),
                    sg=values(r.sg.p),
                    polaris=values(r.p))
summary(v.all)
cor.all <- cor(v.all, use="pairwise.complete.obs")
cor.upper <- cor.all; cor.upper[lower.tri(cor.upper)] <- NA
print(round(cor.upper, 3))
```

Make a nice correlation plot:

```{r corrplot, fig.width=(n.figs.col*1.5), fig.height=n.figs.col*1.5}
library(corrplot)
corrplot(cor.all, method="circle", type="upper", diag=FALSE)
```


# Differences 

## Compute all differences

Relative to gSSURGO, as the base product closest to the field soil survey.

```{r}
diff.gssurgo.sg <- r.gssurgo.p - r.sg.p
diff.gssurgo.p <-  r.gssurgo.p - r.p
```

## Statistics

RMSE, ME, RMSE adjusted to ME, for all products compared to gSSURGO:

```{r stats.compare.sg, warning=FALSE}
stats.compare <- data.frame(PSM_product = "", MD = 0, RMSD = 0, RMSD.Adjusted = 0)
rmse <- function(v1, v2) {
  round(sqrt(mean((v1-v2)^2, na.rm=TRUE)),3)
}
me <- function(v1, v2) { 
  round(mean(v1-v2, na.rm=TRUE), 3)
}
rmse.adj <- function(v1, v2) {   # RMSE adjusted for ME (bias)
  me <- mean(v1-v2, na.rm=TRUE)
  v2.adj <- v2 + me
  round(sqrt(mean((v1-v2.adj)^2, na.rm=TRUE)),3)
}
stats.compare[1, ] <- c("SoilGrids250",
                        me(values(r.gssurgo.p),values(r.sg.p)),
                        rmse(values(r.gssurgo.p),values(r.sg.p)),
                        rmse.adj(values(r.gssurgo.p),values(r.sg.p))
)
stats.compare[2, ] <- c("POLARIS",
                        me(values(r.gssurgo.p),values(r.p)),
                        rmse(values(r.gssurgo.p),values(r.p)),
                        rmse.adj(values(r.gssurgo.p),values(r.p))
)
```

Save this table for incorporation in a LaTeX document:

```{r save.stats.compare.sg}
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
x <- xtable(stats.compare, row.names=FALSE, digits=3)
autoformat(x)
capture.output(print(x, include.rownames=FALSE), 
               file=paste0("../LaTeX_tables/gSSURGO_compare_statistics_",
                           AOI.dir.prefix, "_", voi.sg, "_", depth.list.sg[depth], ".tex"))
```

## Compute common range for all differences

```{r zlim.diff.sg}
zlim <- c(NA, NA)
zlim <- c(floor(min(zlim[1]*10, values(diff.gssurgo.sg)*10, na.rm=TRUE)),
          ceiling(max(zlim[2]*10, values(diff.gssurgo.sg)*10, na.rm=TRUE)))/10
zlim <- c(floor(min(zlim[1]*10, values(diff.gssurgo.p)*10, na.rm=TRUE)),
          ceiling(max(zlim[2]*10, values(diff.gssurgo.p)*10, na.rm=TRUE)))/10
print(zlim)
```

## Histograms

First, compute the maximum density, to have a common y-axis:

```{r diff.hist.densities}
yl <- c(0, max(max.dens(diff.gssurgo.p), max.dens(diff.gssurgo.sg)))
```

Use this y-axis to show all the histograms together:


```{r hist.diff.sg, fig.width=map.fig.width.diff, fig.height=map.fig.height.diff}
par(mfrow=c(n.figs.row.diff, n.figs.col.diff))
hist(diff.gssurgo.p, main="gSSURGO - POLARIS", xlab="",
     xlim=zlim,  breaks=24, freq = FALSE, ylim=yl)
hist(diff.gssurgo.sg, main="gSSURGO - SoilGrids250", xlab="",
     xlim=zlim,  breaks=24, freq = FALSE, ylim=yl)
par(mfrow=c(1,1))
```


## Maps

```{r plot.diff.sg, fig.width=map.fig.width.diff, fig.height=map.fig.height.diff}
par(mfrow=c(n.figs.row.diff, n.figs.col.diff))
terra::plot(diff.gssurgo.p, main="Difference gSSURGO - POLARIS",
            range=zlim, col=bpy.colors(64))
terra::plot(diff.gssurgo.sg, main="Difference gSSURGO - SoilGrids250",
            range=zlim, col=bpy.colors(64))
par(mfrow=c(1,1))
```

# Save harmonized maps {#save}

Set up a directory for these, based on the AOI:

```{r save.dir}
dest.dir.save <-  file.path(base.dir.export,
                       AOI.dir.prefix)
if (!dir.exists(dest.dir.save)) {
   dir.create(dest.dir.save, recursive = TRUE)
}
```

Same AOI, CRS, resolution, units of measure:

```{r save.tiles}
# gSSURGO
voi.depth.sg <- paste0(voi.sg, "_", depth.list.sg[depth])
dest.name <- paste0(dest.dir.save,"/gssurgo_tile_30_",  voi.depth.sg, ".tif")
f <- terra::writeRaster(r.gssurgo.p, file=dest.name,
                        overwrite=TRUE,
                        wopt=list(gdal=c("TFW=YES"), datatype="FLT4S"),
                        filetype="GTiff")
# GDALinfo(dest.name)
print(paste("Wrote ", dest.name))
# POLARIS
dest.name <- paste0(dest.dir.save,"/polaris_tile_30_",  voi.depth.sg, ".tif")
f <- terra::writeRaster(r.p, file=dest.name,
                        overwrite=TRUE,
                        wopt=list(gdal=c("TFW=YES"), datatype="FLT4S"),
                        filetype="GTiff")
#  GDALinfo(dest.name)
print(paste("Wrote ", dest.name))
# SoilGrids250
dest.name <- paste0(dest.dir.save,"/sg_tile_30_",  voi.depth.sg, ".tif")
f <- terra::writeRaster(r.sg.p, file=dest.name,
                        overwrite=TRUE,
                        wopt=list(gdal=c("TFW=YES"), datatype="FLT4S"),
                        filetype="GTiff")
# GDALinfo(dest.name)
print(paste("Wrote ", dest.name))
```
